Single-cell virus sequencing of influenza infections that trigger innate immunity

### 2018-11-15 paper/paper.tex
converted Steuerman re-analysis to supp fig

@@

\subsection{A system to identify and enrich for rare IFN+ positive cells}
A challenge in studying IFN induction by influenza virus is the rareness with which infection triggers its expression.
We previously performed single-cell transcriptomics on hundreds of influenza-infected A549 human lung epithelial cells at early times (6 to 10 hours) post-infection~\citep{russell2018extreme}, and found that only 1 of 368 infected cells expressed detectable type I or type III IFN (\FIG{IFNrare}A).
Prior studies using reporters have found that IFN expression is also rare in influenza-infected mice at 24 and even 48 hours post-infection~\citep{kallfass2013visualizing}.

@@

>>>>

\newfloat{suppfig}{thp}{lofsupfile}
\floatname{suppfig}{Supplementary Figure}

@@

\subsection{A system to identify and enrich for rare IFN+ positive cells}
A challenge in studying IFN induction by influenza virus is the rareness with which infection triggers its expression.
We previously performed single-cell transcriptomics on influenza-infected A549 human lung epithelial cells, and found that only 1 of 368 infected cells expressed detectable type I or type III IFN~\citep{russell2018extreme}.
Prior studies using reporters have found that IFN expression is also rare in influenza-infected cells~\citep{killip2017single} and mice~\citep{kallfass2013visualizing}, a conclusion supported by our own re-analysis of \citet{steuerman2018dissection}'s single-cell transcriptomics of influenza-infected mice (\SUPPFIG{mice}).

@@

\begin{suppfig}
\includegraphics[width=\textwidth]{figures/IFN_stochastic/SteuermanReanalysis/plots/p_mRNA_counts.pdf}
\caption{
A re-analysis of the data from \citet{steuerman2018dissection}'s single-cell transcriptomics of cells from influenza-infected mice shows that only 5 of 1220 virus-infected cells express \emph{detectable} type I or type III IFN transcripts \textit{in vivo}.
Specifically, we downloaded the data from \citet{steuerman2018dissection} and identified influenza-infected cells using criteria similar to those described in \citet{steuerman2018dissection}.
Here we show statistics for the cells from the influenza-infected wild-type C57BL/6J mice, which were collected at 48 hours (two replicates) or 72 hours (one replicate)---we do not show cells from the control mice.
As shown in the left-most panel, the sequencing depth of \citet{steuerman2018dissection} was quite low, with only $\sim$1,500 mRNA counts per cell on average (this is 10 to 15-fold lower than the sequencing depth in \citet{russell2018extreme} and the current study, respectively).
An important caveat is that this low sequencing depth could lead to a simple failure to detect IFN transcripts in some cells.
Nonetheless, there is detectable expression of key ISGs---which are induced by autocrine and paracrine signaling~\citep{stetson2006type,honda2006type}---in the majority of infected cells (the middle panel shows the total counts of IFIT1, ISG15, CCL5, and Mx1).
However, only 5 of 1220 cells express any detectable type I (IFN-$\alpha$ and IFN-$\beta$) or type III (IFN-$\lambda$) mRNAs, which are only induced by direct cellular detection of viral infection~\citep{stetson2006type,honda2006type}.
The full code that performs the re-analysis shown in this figure is at \url{https://github.com/jbloomlab/IFNsorted_flu_single_cell/tree/master/paper/figures/IFN_stochastic/SteuermanReanalysis/}.
}
\label{suppfig:mice}
\end{suppfig}
